---
title: "Project 2 Report"
author: "Tim Cook, Alex Poplawski, and T.J. Seislove"
date: "4/14/2026"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(janitor)
library(lubridate)
library(knitr)
```

# 2020 NFl Analysis

### Report By: Tim Cook, Alex Poplawski, and T.J. Seislove

### Introduction

The dataset for this project is a record of all games from the 2020 NFL season. There are 270 observations, with each representing a game from the 2020 NFL season.

Each row records the week of play, day of the week, date, and the time of the game on the east coast. It also tracks the winner and loser of each game, as well as the numbers of points scored, yards gained, and turnovers committed by each team.

**Dataset Link**

https://www.pro-football-reference.com/years/2020/games.htm

### Data Import

The dataset was downloaded as a csv file and imported using read_csv(). Janitor::clean_names() was used to take care of column names with capitalization and the "/" character. Then, the names of 5 columns were manually changed for clarity using the rename() function and 2 unnecessary columns were dropped with select(). Finally, the date column was converted from characters to date objects using ymd().

```{r}
nfl_2020 <- read_csv("nfl2020.csv") %>% clean_names()

nfl_2020 <- nfl_2020 %>% rename(
  date = date_3,
  pts_w = pts_9,
  pts_l = pts_10,
  to_w = tow,
  to_l = tol
)

nfl_2020 <- nfl_2020 %>% select(-c(x6, date_8))

nfl_2020 <- nfl_2020 %>% mutate(
  date = ymd(date)
)
```

### Exploratory Data Analysis

```{r}
nfl <- read.csv("nfl2020(tidied).csv")

# Total Points
nfl$Total_Pts <- nfl$PtsW + nfl$PtsL
hist(nfl$Total_Pts)
```

This histogram displays the distribution of total points scored in NFL games, the total score was calculated by adding the score of the winning and losing team for each entry. We see that the total scores follow a relatively normal distribution centered around 50 total points.

```{r}
# Total Points vs. Week
plot(nfl$Week, nfl$Total_Pts)
```

This plot takes the same total points metric used above, and displays that data across the 17 weeks. There are no clear patterns across the weeks.

```{r}
# Average points by Month
nfl$Date <- as.Date(nfl$Date)
nfl$Month <- months(nfl$Date)
aggregate(Total_Pts ~ Month, data = nfl, mean) %>% kable()
```

This table displays the average total points scored for all games in a given month of the NFL season. February is notably lower than the rest of the months, but there are no other visible patterns.

### Conclusion

We created two visual summaries, a histogram of NFL total points and a scatter plot total points against week. The histogram illustrates a fairly normal distribution of points scored in games with the highest peak of the scores falling between 40 and 60 and the highest concentration around 50 points. The distribution also suggests that most games had average scoring. However, there were a few games with higher or lower scores indicating that some games had great team performance or poor team performance leading to the difference in scores. Our scatter plot reveals a pretty consistent scoring pattern across the weeks of the NFL season. Looking at the data, scores generally range from 15 to 88 points, with a larger concentration of the points being in between 40 to 60 point range. With the variance being pretty constant, our scatter plot illustrates that NFL teams seemingly maintained a stable performance throughout the football season with our histogram further supporting this.

We created a frequency table portraying the number of games played on each day of the week. Sunday had the highest number of games with 221, which was notably larger than the other days of the week. Most NFL football games are held on Sunday so this number aligns with the traditional NFL schedule. Monday, Thursday, and Saturday followed with a smaller portion of games ranging from 10 to 20 games likely due to Monday and Thursday Night Football and special cases like holidays. Then, the lowest game days were Friday and Wednesday with 1 game each, which can be seen as rest days for the players. Even though Sunday Night Football is the most popular, NFL games occurring on Monday, Thursday, and Saturday is a  strategic decision by the NFL to maximize fan engagement. Additionally, we analyzed the average points scored for games that occurred on the weekday compared to the weekend. We found that average points scored on weekdays is 49.76, which is slightly higher than on weekends which was 49.45. Even though our frequency table shows that most games are played on the weekends, game play factors could influence a teams plan leading to a lower team score the weekday average point statistic shows
 
The monthly average of points illustrates the flow of scoring throughout the NFL season. September, the start of the season, shows the highest monthly points scored. But as the season continues, an increase in competition, pressures regarding playoffs, and injuries could influence scoring dynamics which can cause the point averages to drop. October, December, and February showed slight declines but January had the most substantial drop, from 48.69 in December to 40.00. Understanding these trends can provide insights to measure team performance as the season goes on compared to other teams and allow for strategic planning for future performances. 

